Re-run, Repeat, Reproduce, Reuse, Replicate:\\Transforming Code into Scientific Contributions

### 2017-10-19 draft.tex
last spelling corrections

But precisely because it is easy to believe that if a program runs once and
gives the expected results it will do so forever, crucial steps to transform
working code into meaningful scientific contributions are rarely undertaken
\citep{Collberg:2016,Sandve:2013,Schwab:2000}. Computational research is plagued by replication
problems, in part, because it seems impervious to them. Contrary to production software who provides a service geared towards a practical outcome, the motivation behind scientific code is to test an hypothesis. While in some instance production software and scientific code are indistinguishable, the reasons why they were created are different, and, therefore, so are the criteria to evaluate their success. A program can
fail as a scientific contribution in many different ways for many different
reasons. Borrowing the terms coined by \citeauthor{Goble:2016}
\citep{Goble:2016}, for a program to contribute to science, it should be
re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$),
reusable (R$^4$) and replicable (R$^5$). Let us illustrate this with a small
example, a random walk \citep{Hughes:1995} written in Python:\\

@@

The code is running and producing the expected results. The next step is to make sure that you can produce the same output over successive runs of your program. In other words, the next step is to make your program deterministic, producing {\em repeatable} output.  Repeatability is valuable. If a run of the program produces a particularly puzzling result, repeatability allows you to scrutinize any step of the execution of the program by re-running it again with extraneous prints, or inside a debugger. Repeatability is also useful to prove that the program did indeed produce the published results. Repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}. But for sequential and deterministically parallels programs \citep{Hines:2008, Collange:2015} not depending on analog inputs, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\

@@

% reproducible, a definition
\emph{Replicability} is the implicit
assumption that an article that does not provide the source code makes: that
the description it provides of the algorithms is sufficiently precise and
complete to re-obtain the results it presents. Here, replicating implies writing a new code matching the conceptual description of the article, in order to reobtain the same results. Replication affords robustness to the results because, should the original code contain an error, a different codebase creates the possibility that this error will not be repeated, in the same way that replicating an laboratory experiment in a different laboratory can ferret out subtle biases. While every published article should strive for replicability, it is seldom obtained. In fact, absent an explicit effort to make an algorithmic description replicable, there is little probability that it will be.\\

>>>>

But precisely because it is easy to believe that if a program runs once and
gives the expected results it will do so forever, crucial steps to transform
working code into meaningful scientific contributions are rarely undertaken
\citep{Collberg:2016,Sandve:2013,Schwab:2000}. Computational research is plagued by replication
problems, in part, because it seems impervious to them. Contrary to production software who provides a service geared towards a practical outcome, the motivation behind scientific code is to test a hypothesis. While in some instance production software and scientific code are indistinguishable, the reasons why they were created are different, and, therefore, so are the criteria to evaluate their success. A program can
fail as a scientific contribution in many different ways for many different
reasons. Borrowing the terms coined by \citeauthor{Goble:2016}
\citep{Goble:2016}, for a program to contribute to science, it should be
re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$),
reusable (R$^4$) and replicable (R$^5$). Let us illustrate this with a small
example, a random walk \citep{Hughes:1995} written in Python:\\

@@

The code is running and producing the expected results. The next step is to make sure that you can produce the same output over successive runs of your program. In other words, the next step is to make your program deterministic, producing {\em repeatable} output.  Repeatability is valuable. If a run of the program produces a particularly puzzling result, repeatability allows you to scrutinize any step of the execution of the program by re-running it again with extraneous prints, or inside a debugger. Repeatability is also useful to prove that the program did indeed produce the published results. Repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}. But for sequential and deterministically parallel programs \citep{Hines:2008, Collange:2015} not depending on analog inputs, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\

@@

% reproducible, a definition
\emph{Replicability} is the implicit
assumption that an article that does not provide the source code makes: that
the description it provides of the algorithms is sufficiently precise and
complete to re-obtain the results it presents. Here, replicating implies writing a new code matching the conceptual description of the article, in order to reobtain the same results. Replication affords robustness to the results because, should the original code contain an error, a different codebase creates the possibility that this error will not be repeated, in the same way that replicating a laboratory experiment in a different laboratory can ferret out subtle biases. While every published article should strive for replicability, it is seldom obtained. In fact, absent an explicit effort to make an algorithmic description replicable, there is little probability that it will be.\\
